\Ace: Growing a Statically-Typed Language Inside a Python

### 2013-11-15 ace-pldi14/ace-pldi14.tex
pldi14 merged heathers changes

@@

Generic functions are represented at compile-time as instances of \verb|ace.GenericFn| and consist of an abstract syntax tree, an {active base} and an {active target}. The purpose of the \emph{decorator} on line 3 is to replace the Python function on lines 4-8 with an Ace generic function having the same syntax tree and the provided active base and active target. 
Decorators in Python are simply syntactic sugar for applying the decorator function directly to the function  being decorated \cite{python}. In other words, line 3 could be replaced by inserting the following  statement on line 9\todo{figure out how to center this}:

\verb|map = ace.fn(clx.base, clx.opencl)(map)|

@@

\section{Related Work}\label{related}
\begin{figure*}
\begin{tabular}[t]{ l | l | c c c c c }
\textbf{Approach} & \textbf{Examples} & Library & {Extensible} & Extensible & Composable & Alternative \\
\hline
Active Types & Ace & x & ~ & x & x & x\\
Desugaring & SugarJ, Sugar* & ~ & x & ~ & ~ & ~\\
Rule Injection & Qi, Typed Racket/Clojure, A? & 1+2 & ~ & x & ~ & ~\\
Static metaprogramming & Scala macros, OJ, OC++ & ~ & ~ & ~ & x & ~\\
Cross-compilation & Delite & x & ~ & ~ & ~ & x\\
EDSL frameworks & ? & x & x & x & ~ & ~\\
Type-specific literals & Wyvern & ~ & x & ~ & x & ~\\
\end{tabular}
\caption{Comparison to related approaches}
\end{figure*}
\subsection{Active Libraries}
Libraries that contain compile-time logic have been called {\it active libraries} in prior proposals \cite{activelibraries}. A number of  projects, such as Blitz++, have taken advantage of the C++ preprocessor and template-based metaprogramming system to implement domain-specific optimizations \cite{veldhuizen2000blitz++}. In Ace, we replace these brittle mini-languages with a general-purpose language and significantly expand the notion of active libraries by consideration of types as objects in the metalanguage. We thus call these types \emph{active types}.

@@

 Future work: integrate with something like scalad http://lampwww.epfl.ch/~hmiller/scala2013/resources/pdfs/paper8.pdf.

@@

\bibliographystyle{abbrvnat}

>>>>

 % Heather-added packages for the fancy table
 \usepackage{longtable}
 \usepackage{booktabs}
 \usepackage{pdflscape}
 \usepackage{colortbl}%
 \newcommand{\myrowcolour}{\rowcolor[gray]{0.925}}
 \usepackage{wasysym}

@@

Generic functions are represented at compile-time as instances of \verb|ace.GenericFn| and consist of an abstract syntax tree, an {active base} and an {active target}. The purpose of the \emph{decorator} on line 3 is to replace the Python function on lines 4-8 with an Ace generic function having the same syntax tree and the provided active base and active target. 
Decorators in Python are simply syntactic sugar for applying the decorator function directly to the function  being decorated \cite{python}. In other words, line 3 could be replaced by inserting the following  statement on line 9:
\vspace{-0.2cm}
\begin{verbbox}
map = ace.fn(clx.base, clx.opencl)(map)
\end{verbbox}
\begin{figure}[h!]
\centering
\theverbbox
\end{figure}
\vspace{-0.2cm}

@@

\section{Related Work}\label{related}
\parindent0pt
%
\begin{figure*}
\onecolumn
\begin{longtable}{l l@{}l c@{}c c@{}c c@{}c c@{}c c@{}c}
%\toprule%

&&  &&  && {\bfseries Extensible} && {\bfseries Extensible} &&  && {\bfseries Alternative}\\

{\bfseries Approach} && {\bfseries Examples} && {\bfseries Library} && {\bfseries Syntax} && {\bfseries Type System} && {\bfseries Compositional} && {\bfseries Targets}  \\

\cmidrule(l){1-1} \cmidrule(l){2-3} \cmidrule(l){4-5} \cmidrule(l){6-7} \cmidrule(l){8-9} \cmidrule(l){10-11} \cmidrule(l){12-13}

\endhead

Active Types && Ace && \CIRCLE && \Circle && \CIRCLE && \CIRCLE && \CIRCLE \\

\myrowcolour%
Desugaring && SugarJ \cite{erdweg2011sugarj}, Sugar* && \Circle && \CIRCLE && \Circle && \Circle && \Circle \\

Rule Injection && Qi \cite{qi}, Typed Racket \cite{TypedScheme2008}, A? && 1+2 && \Circle && \CIRCLE && \Circle && \Circle \\

\myrowcolour%
Static macros /  && Scala macros \cite{ScalaMacros2013}, MorphJ \cite{MorphJ2011}, && \Circle && \Circle && \Circle && \CIRCLE && \Circle \\

\myrowcolour
Metaprogramming &&  OJ \cite{OpenJava2000}, Template Haskell \cite{template-haskell} && && && && && \\

Cross-Compilation && Delite \cite{Delite2011} && \CIRCLE && \Circle && \Circle && \Circle && \CIRCLE \\

\myrowcolour%
EDSL Frameworks && ? && \CIRCLE && \CIRCLE && \CIRCLE && \Circle && \Circle \\

Type-Specific Literals && Wyvern \cite{globaldsl13} && \Circle && \CIRCLE && \Circle && \CIRCLE && \Circle \\
\end{longtable}
\caption{Comparison to related approaches}
\twocolumn
\end{figure*}
\subsection{Active Libraries}
Libraries that contain compile-time logic have been called {\it active libraries} in prior proposals \cite{activelibraries}. A number of  projects, such as Blitz++, have taken advantage of the C++ preprocessor and template-based metaprogramming system to implement domain-specific optimizations \cite{veldhuizen2000blitz++}. In Ace, we replace these brittle mini-languages with a general-purpose language and significantly expand the notion of active libraries by consideration of types as objects in the metalanguage. We thus call these types \emph{active types}.

@@

 Future work: integrate with something like scalad http://lampwww.epfl.ch/~hmiller/scala2013/resources/pdfs/paper8.pdf. \todo{fix python lang reference citation}

@@

\bibliographystyle{abbrv}
